---
title: "L09-Assignment"
output: html_notebook
---

# Assignment Instructions

* Complete all cells as instructed, replacing any ??? with the appropriate code

* Execute RStudio **Session** > **Restart & Run All Chunks** and ensure that all code blocks run without error

* Do not load any additional libraries. You must use the libraries in the Load Libraries cell provided.

* Meet all of the assignment objectives described below

* RStudio Notebook cells can be added and inserted as desired

# Assignment Objectives


Classify iris flower species. This includes:

* 1 point - All code blocks run without error. RStudio **Session** > **Restart & Run All Chunks**
* 1 point - Create 2 classification models, with at least one multivariate classification model
* 1 point - Confusion matrix for each model
* 1 point - Chart visualization for each model
* 1 point - Training set and testing set for each model
* 1 point - Include the following R Features at least once. You are not limited to these, only by the libraries loaded.
* 1 point - Include a written analysis of iris species category performance referencing the confusion matrix as evidence of your conclusions.

The above is what is required to achieve full credit for the assignment. You are welcome and encouraged to go above and beyond these requirements, just be sure these requirements are fully met first. 


## R Features
* caret::confusionMatrix()
* set.seed()
* glm()
* predict()

## Dataset
* iris

```{r}
# Load libraries
# Do not load any additional libraries other than what is provided in this template

# You can reference the caret package using caret::<function>
library(lubridate)
library(tidyverse)
```

```{r}
# Explore data structure
# Data: iris
iris %>% glimpse()

# Display help on data
? iris
```

# Data processing
Create a new data frame(s) with appropriate data types and data cleaning for the data.



```{r}
#Added this line to make the data a binary classification
df.iris <- iris %>% mutate(Species=if_else(Species=="versicolor", TRUE, FALSE) %>% as.factor())
# Placeholder cell to create your own data frames from the source data
normalize <- function(x) (x - mean(x))/sd(x)
df.iris  = df.iris %>%
    mutate(Sepal.Length.Scaled = normalize(Sepal.Length),
           Sepal.Width.Scaled = normalize(Sepal.Width),
           Petal.Length.Scaled = normalize(Petal.Length),
           Petal.Width.Scaled = normalize(Petal.Width)
           )
df.iris %>% glimpse() 
df.iris %>% summary()
```
```{r}
## 80% of the sample size
set.seed(1222)
iris.train <- df.iris %>% 
  sample_frac(0.8)

iris.test <- df.iris %>%  
  setdiff(iris.train)

print(str_c("iris.train rows: ", nrow(iris.train)))
print(str_c("iris.test rows: ", nrow(iris.test)))
nrow(iris.test) + nrow(iris.train) == nrow(df.iris)

iris.test$Species %>% table()
```

# Classification Model 1
The model has small p-values and z-scores, the model may be an overfit. However, the prediction accuarcy, precision, and recall are all somewhat low.  
```{r}
# placeholder
iris.train %>% 
  ggplot(aes(x = Species, y = Sepal.Length.Scaled)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75), 
              fill = 'blue', alpha = 0.3, size = 1.0) +
  labs(title= "Species  by  Sepal.Length")

Sepal.Length.mod <- glm(Species ~ Sepal.Length.Scaled, 
                   data = iris.train, 
                family=binomial())
Sepal.Length.mod %>% summary()

cat('Confidence intervals of coefficients')
Sepal.Length.mod %>% confint()
```

```{r}
iris.test$score <- predict(Sepal.Length.mod, newdata = iris.test)
binwidth <- (max(iris.test[, 'score']) - min(iris.test[, 'score'])) / 60

ggplot(iris.test, aes(x = score)) + 
  geom_dotplot(dotsize = 0.5, method = "histodot", binwidth = binwidth) +
  labs(title="Model Score by True or False")

iris.test = iris.test %>% 
                    mutate(score.species.prob = exp(score)/(1 + exp(score))) 
iris.test %>% glimpse()

iris.test %>% 
  ggplot(aes(x = Species, y = score.species.prob)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75), 
              fill = 'blue', alpha = 0.3, size = 1.0) +
  labs(title= "Species  by  Sepal.Length")
```


```{r}
iris.test <- iris.test %>% mutate(score.species = if_else(score.species.prob > 0.32, "TRUE", "FALSE") %>% as.factor())

iris.test %>% glimpse()

iris.test$score.species %>% table()
caret::confusionMatrix(data = iris.test$score.species, 
                       reference = iris.test$Species, 
                       mode = "prec_recall")
```


# Classification Model 2
The z-score and p-value are small, and the model may be an overfit. The accruacy, precision and recall values are somewhat acceptable.
```{r}
plot.violin <- function(col, df){
    p1 <- df %>% 
      ggplot(aes_string('Species', col)) + 
      geom_violin(draw_quantiles = c(0.25, 0.5, 0.75), 
              fill = 'blue', alpha = 0.3, size = 1.0)
    p1 %>% print()
}
cols = df.iris %>% 
        select(Sepal.Length.Scaled, Sepal.Width.Scaled, Petal.Length.Scaled, Petal.Width.Scaled)

cols %>% walk(plot.violin, df.iris)

```

```{r}
Species.mod <- glm(Species ~., 
                   data = iris.train, 
                family=binomial())

Species.mod %>% summary()

cat('Confidence intervals of coefficients')
Species.mod %>% confint()
```
```{r}
iris.test$score2 <- predict(Species.mod, newdata = iris.test)
binwidth <- (max(iris.test[, 'score2']) - min(iris.test[, 'score2'])) / 60

ggplot(iris.test, aes(x = score2)) + 
  geom_dotplot(dotsize = 0.5, method = "histodot", binwidth = binwidth) +
  labs(title="Model Score by True or False")

iris.test = iris.test %>% 
                    mutate(score.species.prob2 = exp(score2)/(1 + exp(score2))) 
iris.test %>% glimpse()

iris.test %>% 
  ggplot(aes(x = Species, y = score.species.prob2)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75), 
              fill = 'blue', alpha = 0.3, size = 1.0) +
  labs(title= "Species  by  Sepal & Petal")
```
```{r}
iris.test <- iris.test %>% mutate(score.species2 = if_else(score.species.prob2 > 0.37, "TRUE", "FALSE") %>% as.factor())

iris.test %>% glimpse()

iris.test$score.species2 %>% table()
caret::confusionMatrix(data = iris.test$score.species2, 
                       reference = iris.test$Species, 
                       mode = "prec_recall")
```



# Analysis and Conclusions
The first model predicts the species by sepal length whether it is versicolor(true) or false. The first model gives a somewhat in accruate model since the accuracy, precision and recall values of the prediction are fairly low. The second model predicts the species by sepal and pedal length and width to determine the species. The accruacy, precision and recall values are somewhat acceptable.
